---
title: "Univariate statistical analyses to identify diffrentially abundant proteins."
author: "Ashfaq Ali"
date: "30/11/2021"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE, 
                      collapse = TRUE, 
                      warning = FALSE, 
                      message = FALSE, 
                      cache = FALSE, 
                      cache.path = "../cache/", 
                      fig.height = 5, 
                      fig.width = 6,
                      eval=TRUE)
```

```{r setup1, echo=FALSE ,include=FALSE}
# import libraries and data

library(tidyverse)
library(reshape2)
library(readxl)
library(fuzzyjoin)
library(limma)
library(DBI)
library(ggplot2)
#library(BiocGenerics)
#library(devtools)
#library(qvalue)
library(pheatmap)
library(RColorBrewer)
library(SummarizedExperiment)
```

# Summary


This document contains analyses code to identify differential abundant proteins using uni variate approach where `limma()` package in R was used to compare abundance differences for following grouping variables.

As batch effect was observed in the in the data and analyses between different time-points maybe confounded by the batch, protein data from the 3 time-points was analysed separately.

1) Temperature (Randomisation code)
2) cpc score (dichotamised)
3) Temperature:CPC score interaction
4) Shockable vs. non-shockable heart


## Import data and perform statistics

Protein annotations were requested 

```{r, echo=FALSE, include=FALSE}
data_all <- readRDS("../rawaDataSe_fixedIDs.rds")
#data_all <- readRDS("../normalized_filtered_se.rds")
data_all <- data_all[, !is.na(data_all$randomisation_code)]
rowData(data_all)$variance <- assays(data_all)$loess %>% rowVars(na.rm = TRUE)
#data_all <- data_all[rowData(data_all)$variance>0.2 | rowData(data_all)$variance<5, ]
data_24h <- data_all[, data_all$time=="-24-"]
data_48h <- data_all[, data_all$time=="-48-"]
data_72h <- data_all[, data_all$time=="-72-"]
data <- list(loess_data= assays(data_all)$loess %>% as.data.frame() %>% rownames_to_column(., var = "Protein.Group") , sample_data=colData(data_all) %>% as.data.frame() %>% rownames_to_column(., var = "Sample_ID"), proetin_anno=as.data.frame(rowData(data_all)))
writexl::write_xlsx(x = data , "./Results/Data_v2.xlsx")
```


## 1) Temperature Randomization code

```{r, eval=FALSE, include=FALSE}


#designAll <- model.matrix(~ 1 + data_all$randomisation_code)
design24h <- model.matrix(~ 1 + data_24h$randomisation_code)
design48h <- model.matrix(~ 1 + data_48h$randomisation_code)
design72h <- model.matrix(~ 1 + data_72h$randomisation_code)

#colnames(designAll) <- c("(Intercept)", "rand_cod")
colnames(design24h) <- c("(Intercept)", "rand_cod")
colnames(design48h) <- c("(Intercept)", "rand_cod")
colnames(design72h) <- c("(Intercept)", "rand_cod")


Results <- list()

# fitAll <- lmFit((assays(data_all)$loess), design = designAll)
# fitAll <- eBayes(fit = fitAll)
# Results$All_rand <- topTable(fitAll, p.value = 1, number = Inf, coef = "rand_cod")

fit24<- lmFit((assays(data_24h)$loess), design = design24h) 

fit24 <- eBayes(fit = fit24)
Results$rand24 <- topTable(fit24, p.value = 1, number = Inf, coef = "rand_cod") %>%
  as.data.frame() %>%
  rownames_to_column(., var = "ID") %>%
 # arrange(padj) %>%
  right_join(rowData(data_all) |> as_tibble(),
    {
      .
    },
    by ="ID"
  ) 
    
    
fit48 <- lmFit((assays(data_48h)$loess), design = design48h)
fit48 <- eBayes(fit = fit48)
Results$rand_48 <- topTable(fit48, p.value = 1, number = Inf, coef = "rand_cod") %>%
  rownames_to_column(., var = "ID") %>%
 # arrange(padj) %>%
  right_join(rowData(data_all) |> as_tibble(),
    {
      .
    },
    by ="ID"
  ) 

fit72 <- lmFit((assays(data_72h)$loess), design = design72h)
fit72 <- eBayes(fit = fit72)
Results$rand72 <- topTable(fit72, p.value = 1, number = Inf, coef = "rand_cod") %>%
  rownames_to_column(., var = "ID") %>%
 # arrange(padj) %>%
  right_join(rowData(data_all) |> as_tibble(),
    {
      .
    },
    by ="ID"
  ) 

```

```{r, eval=FALSE, include=FALSE}
batches <- unique(data_all$time)
Results_rand_v2 <- list()
Results_dead72_v2 <- list()
Results_folllow_up_v2 <- list()
Results_cpc_v2 <- list()
Results_shock_v2 <- list()
Results_shock_int_v2 <- list()
Results_cpc_int_v2 <- list()



for  (i in 1:length(batches)){
  data <- data_all[, data_all$time == batches[i]] 
  data_shock <- data[, data$shockable!="NA"] 
  #data <- season_batches[[batches[i]]]

### Randomization code    
  design_t <- model.matrix(~ 1 + randomisation_code, data = colData(data))
  colnames(design_t) <- c("(Intercept)", "rand_cod")
  
fit<- lmFit((assays(data)$loess), design = design_t) 

fit <- eBayes(fit = fit)

Results_rand_v2[[batches[i]]] <- topTable(fit, p.value = 1, number = Inf, coef = "rand_cod", confint = TRUE) %>%
  as.data.frame() %>%
  rownames_to_column(., var = "ID") %>%
 # arrange(padj) %>%
  left_join(rowData(data) |> as_tibble(),
    {
      .
    },
    by ="ID"
  ) 
}
writexl::write_xlsx(Results_rand_v2, paste0("./Results/DE_results_rand_code_v2.xlsx"))
saveRDS(Results_rand_v2, "./Results/DE_Results_rand_v2.rds")




################# CPC dead at 72h  ############

for  (i in 1:length(batches)){
  data <- data_all[, data_all$time == batches[i]] 
  
design_d_72 <- model.matrix(~ 1 + dead72, data = colData(data))
colnames(design_d_72) <- c("(Intercept)", "dead_72")

  fit<- lmFit((assays(data)$loess), design = design_d_72) 

fit <- eBayes(fit = fit)

Results_dead72_v2[[batches[i]]] <- topTable(fit, p.value = 1, number = Inf, coef = "dead_72", confint = TRUE) %>%
  as.data.frame() %>%
  rownames_to_column(., var = "ID") %>%
 # arrange(padj) %>%
  left_join(rowData(data) |> as_tibble(),
    {
      .
    },
    by ="ID"
  ) }
writexl::write_xlsx(Results_dead72_v2, paste0("./Results/DE_results_dead_at_72_v2.xlsx"))
saveRDS(Results_dead72_v2, "./Results/DE_Results_dead_at_72_v2.rds")


################# CPC dead at follow up  ############
for  (i in 1:length(batches)){
   data <- data_all[, data_all$time == batches[i]] 
design_d_fup <- model.matrix(~ 1 + deadatfollowup, data = colData(data))
colnames(design_d_fup) <- c("(Intercept)", "dead_fup")

  fit<- lmFit((assays(data)$loess), design = design_d_fup) 

fit <- eBayes(fit = fit)

  Results_folllow_up_v2[[batches[i]]] <- topTable(fit, p.value = 1, number = Inf, coef = "dead_fup", confint = TRUE) %>%
  as.data.frame() %>%
  rownames_to_column(., var = "ID") %>%
 # arrange(padj) %>%
  left_join(rowData(data) |> as_tibble(),
    {
      .
    },
    by ="ID"
  ) 
}
writexl::write_xlsx(Results_folllow_up_v2, paste0("./Results/DE_results_dead_at_follow_up_v2.xlsx"))
saveRDS(Results_folllow_up_v2, "./Results/DE_Results_dead_at_follow_up_v2.rds")


################# CPC Score ############

for  (i in 1:length(batches)){
  data <- data_all[, data_all$time == batches[i]] 
design_cpc <- model.matrix(~ 1 + binarysecout180daycpcscore, data = colData(data))
colnames(design_cpc) <- c("(Intercept)", "cpc_score")

fit<- lmFit((assays(data)$loess), design = design_cpc) 

fit <- eBayes(fit = fit)

Results_cpc_v2[[batches[i]]] <- topTable(fit, p.value = 1, number = Inf, coef = "cpc_score", confint = TRUE) %>%
  as.data.frame() %>%
  rownames_to_column(., var = "ID") %>%
 # arrange(padj) %>%
  left_join(rowData(data) |> as_tibble(),
    {
      .
    },
    by ="ID"
  ) 
}
writexl::write_xlsx(Results_cpc_v2, paste0("./Results/DE_results_cpc_score_v2.xlsx"))
saveRDS(Results_cpc_v2, "./Results/DE_Results_cpc_v2.rds")


##############3 Shock variable ###########
## 

for  (i in 1:length(batches)){
  data <- data_all[, data_all$time == batches[i]] 
  
design_shock <- model.matrix(~ 1 + shockable, data = colData(data_shock))
colnames(design_shock) <- c("(Intercept)", "shock")
fit<- lmFit((assays(data_shock)$loess), design = design_shock) 

fit <- eBayes(fit = fit)

Results_shock_v2[[batches[i]]] <- topTable(fit, p.value = 1, number = Inf, coef = "shock", confint = TRUE) %>%
  as.data.frame() %>%
  rownames_to_column(., var = "ID") %>%
 # arrange(padj) %>%
  left_join(rowData(data) |> as_tibble(),
    {
      .
    },
    by ="ID"
  ) 
}
writexl::write_xlsx(Results_shock_v2, paste0("./Results/DE_results_shockable_vs_non_shockable_v2.xlsx"))
saveRDS(Results_shock_v2, "./Results/DE_Results_shock_v2.rds")


################## Interaction with CPC #################
## 
for  (i in 1:length(batches)){
  data <- data_all[, data_all$time == batches[i]] 
design_cpc_int <- model.matrix(~ 1 + randomisation_code*binarysecout180daycpcscore, data = colData(data))

# colnames(designAll) <- c("(Intercept)", "rand_cod", "pcscore", "rand_codXpcscore")
colnames(design_cpc_int) <- c("(Intercept)", "rand_cod", "cpc", "rand_codXcpc")
fit<- lmFit((assays(data)$loess), design = design_cpc_int) 

fit <- eBayes(fit = fit)

Results_cpc_int_v2[[batches[i]]] <- topTable(fit, p.value = 1, number = Inf, coef = "rand_codXcpc", confint = TRUE) %>%
  as.data.frame() %>%
  rownames_to_column(., var = "ID") %>%
 # arrange(padj) %>%
  left_join(rowData(data) |> as_tibble(),
    {
      .
    },
    
    by ="ID"
  ) 
}
writexl::write_xlsx(Results_cpc_int_v2, paste0("./Results/DE_results_cpc_rand_Code_int_v2.xlsx"))
saveRDS(Results_cpc_int_v2, "./Results/DE_results_cpc_int_v2.rds")


############## Rand_shock interaction  ##############
for  (i in 1:length(batches)){
  data <- data_all[, data_all$time == batches[i]] 
data_shock <- data[, !(data$shockable=="NA")] 
data_shock$shockable <- droplevels(data_shock$shockable)
design_shock_int <- model.matrix(~ 1 + randomisation_code*shockable, data = colData(data_shock))
colnames(design_shock_int) <- c("(Intercept)", "rand_cod", "shock", "rand_codXshock")
fit<- lmFit((assays(data_shock)$loess), design = design_shock_int) 

fit <- eBayes(fit = fit)

Results_shock_int_v2[[batches[i]]] <- topTable(fit, p.value = 1, number = Inf, coef = "rand_codXshock", confint = TRUE) %>%
  as.data.frame() %>%
  rownames_to_column(., var = "ID") %>%
 # arrange(padj) %>%
  left_join(rowData(data_shock) |> as_tibble(),
    {
      .
    },
    
    by ="ID"
  )
}
writexl::write_xlsx(Results_shock_int_v2, paste0("./Results/DE_results_shock_rand_Code_int_v2.xlsx"))
saveRDS(Results_shock_int_v2, "./Results/DE_results_shock_rand_Code_int_v2.rds")

```
****************************************************


```{r, eval=FALSE, include=FALSE}

# designAll <- model.matrix(~ 1 + data_all$binarysecout180daycpcscore)
design24h <- model.matrix(~ 1 + data_24h$binarysecout180daycpcscore)
design48h <- model.matrix(~ 1 + data_48h$binarysecout180daycpcscore)
design72h <- model.matrix(~ 1 + data_72h$binarysecout180daycpcscore)

#colnames(designAll) <- c("(Intercept)", "pcscore")
colnames(design24h) <- c("(Intercept)", "pcscore")
colnames(design48h) <- c("(Intercept)", "pcscore")
colnames(design72h) <- c("(Intercept)", "pcscore")


#Results2 <- list()

# fitAll <- lmFit((assays(data_all)$loess), design = designAll)
# fitAll <- eBayes(fit = fitAll)
# Results$All_pcscore <- topTable(fitAll, p.value = 1, number = Inf, coef = "pcscore")

fit24<- lmFit((assays(data_24h)$loess), design = design24h)
fit24 <- eBayes(fit = fit24)
Results$pcscore24 <- topTable(fit24, p.value = 1, number = Inf, coef = "pcscore") %>%
  rownames_to_column(., var = "ID") %>%
 # arrange(padj) %>%
  right_join(rowData(data_all) |> as_tibble(),
    {
      .
    },
    by ="ID"
  ) 

fit48 <- lmFit((assays(data_48h)$loess), design = design48h)
fit48 <- eBayes(fit = fit48)
Results$pcscore48 <- topTable(fit48, p.value = 1, number = Inf, coef = "pcscore") %>%
  rownames_to_column(., var = "ID") %>%
 # arrange(padj) %>%
  right_join(rowData(data_all) |> as_tibble(),
    {
      .
    },
    by ="ID"
  ) 

fit72 <- lmFit((assays(data_72h)$loess), design = design72h)
fit72 <- eBayes(fit = fit72)
Results$pcscore72 <- topTable(fit72, p.value = 1, number = Inf, coef = "pcscore") %>%
  rownames_to_column(., var = "ID") %>%
 # arrange(padj) %>%
  right_join(rowData(data_all) |> as_tibble(),
    {
      .
    },
    by ="ID"
  ) 

rm(fit24, fit48h, fit72, design24h, design48h, design72h)
#saveRDS(Results, "../DEResults.rds")
```


### Tables {.tabset}

```{r, eval=FALSE, include=FALSE}
# designAll <- model.matrix(~ 1 + data_all$randomisation_code*data_all$binarysecout180daycpcscore)
design24h <- model.matrix(~ 1 + data_24h$randomisation_code*data_24h$binarysecout180daycpcscore)
design48h <- model.matrix(~ 1 + data_48h$randomisation_code*data_48h$binarysecout180daycpcscore)
design72h <- model.matrix(~ 1 + data_72h$randomisation_code*data_72h$binarysecout180daycpcscore)

# colnames(designAll) <- c("(Intercept)", "rand_cod", "pcscore", "rand_codXpcscore")
colnames(design24h) <- c("(Intercept)", "rand_cod", "pcscore", "rand_codXpcscore")
colnames(design48h) <- c("(Intercept)", "rand_cod", "pcscore", "rand_codXpcscore")
colnames(design72h) <- c("(Intercept)", "rand_cod", "pcscore", "rand_codXpcscore")


#Results2 <- list()
# 
# fitAll <- lmFit((assays(data_all)$loess), design = designAll)
# fitAll <- eBayes(fit = fitAll)
# Results$All_randpcs_int <- topTable(fitAll, p.value = 1, number = Inf, coef = "rand_codXpcscore")

fit24<- lmFit((assays(data_24h)$loess), design = design24h)
fit24 <- eBayes(fit = fit24)
Results$randpcs_int24 <- topTable(fit24, p.value = 1, number = Inf, coef = "rand_codXpcscore") %>%
  rownames_to_column(., var = "ID") %>%
 # arrange(padj) %>%
  right_join(rowData(data_all) |> as_tibble(),
    {
      .
    },
    by ="ID"
  ) 

fit48 <- lmFit((assays(data_48h)$loess), design = design48h)
fit48 <- eBayes(fit = fit48)
Results$randpcs_int48 <- topTable(fit48, p.value = 1, number = Inf, coef = "rand_codXpcscore") %>%
  rownames_to_column(., var = "ID") %>%
 # arrange(padj) %>%
  right_join(rowData(data_all) |> as_tibble(),
    {
      .
    },
    by ="ID"
  ) 

fit72 <- lmFit((assays(data_72h)$loess), design = design72h)
fit72 <- eBayes(fit = fit72)
Results$randpcs_int72 <- topTable(fit72, p.value = 1, number = Inf, coef = "rand_codXpcscore") %>%
  rownames_to_column(., var = "ID") %>%
 # arrange(padj) %>%
  right_join(rowData(data_all) |> as_tibble(),
    {
      .
    },
    by ="ID"
  ) 
#saveRDS(Results, "../NBIS_analyses/Results/DEResults.rds")
```

##4) Dead vs Alive at 72h

No patients were declared dead at 24 or 48 hours, therefor death at 72 hours was used for assications with proteins at 24, 48, and 72 hours.



```{r}

# designAll <- model.matrix(~ 1 + data_all$binarysecout180daycpcscore)
design24h <- model.matrix(~ 1 + data_24h$dead72)
design48h <- model.matrix(~ 1 + data_48h$binarysecout180daycpcscore)
design72h <- model.matrix(~ 1 + data_72h$binarysecout180daycpcscore)

#colnames(designAll) <- c("(Intercept)", "pcscore")
colnames(design24h) <- c("(Intercept)", "pcscore")
colnames(design48h) <- c("(Intercept)", "pcscore")
colnames(design72h) <- c("(Intercept)", "pcscore")


#Results2 <- list()

# fitAll <- lmFit((assays(data_all)$loess), design = designAll)
# fitAll <- eBayes(fit = fitAll)
# Results$All_pcscore <- topTable(fitAll, p.value = 1, number = Inf, coef = "pcscore")

fit24<- lmFit((assays(data_24h)$loess), design = design24h)
fit24 <- eBayes(fit = fit24)
Results$pcscore24 <- topTable(fit24, p.value = 1, number = Inf, coef = "pcscore") %>%
  rownames_to_column(., var = "ID") %>%
 # arrange(padj) %>%
  right_join(rowData(data_all) |> as_tibble(),
    {
      .
    },
    by ="ID"
  ) 

fit48 <- lmFit((assays(data_48h)$loess), design = design48h)
fit48 <- eBayes(fit = fit48)
Results$pcscore48 <- topTable(fit48, p.value = 1, number = Inf, coef = "pcscore") %>%
  rownames_to_column(., var = "ID") %>%
 # arrange(padj) %>%
  right_join(rowData(data_all) |> as_tibble(),
    {
      .
    },
    by ="ID"
  ) 

fit72 <- lmFit((assays(data_72h)$loess), design = design72h)
fit72 <- eBayes(fit = fit72)
Results$pcscore72 <- topTable(fit72, p.value = 1, number = Inf, coef = "pcscore") %>%
  rownames_to_column(., var = "ID") %>%
 # arrange(padj) %>%
  right_join(rowData(data_all) |> as_tibble(),
    {
      .
    },
    by ="ID"
  ) 

rm(fit24, fit48h, fit72, design24h, design48h, design72h)
#saveRDS(Results, "../DEResults.rds")
```

##5) Dead at follow up




##6) Interaction analyses Shockabl*Randomization code


```{r}
# designAll <- model.matrix(~ 1 + data_all$randomisation_code*data_all$binarysecout180daycpcscore)
design24h <- model.matrix(~ 1 + data_24h$randomisation_code*data_24h$shockable)
design48h <- model.matrix(~ 1 + data_48h$randomisation_code*data_48h$shockable)
design72h <- model.matrix(~ 1 + data_72h$randomisation_code*data_72h$shockable)

# colnames(designAll) <- c("(Intercept)", "rand_cod", "pcscore", "rand_codXpcscore")
colnames(design24h) <- c("(Intercept)", "rand_cod", "shock", "rand_codXshock")
colnames(design48h) <- c("(Intercept)", "rand_cod", "shock", "rand_codXshock")
colnames(design72h) <- c("(Intercept)", "rand_cod", "shock", "rand_codXshock")


#Results2 <- list()
# 
# fitAll <- lmFit((assays(data_all)$loess), design = designAll)
# fitAll <- eBayes(fit = fitAll)
# Results$All_randpcs_int <- topTable(fitAll, p.value = 1, number = Inf, coef = "rand_codXpcscore")

fit24 <- lmFit((assays(data_24h)$loess), design = design24h)
fit24 <- eBayes(fit = fit24)
Results$randpcs_int24 <- topTable(fit24, p.value = 1, number = Inf, coef = "rand_codXshock") %>%
  rownames_to_column(., var = "ID") %>%
 # arrange(padj) %>%
  right_join(rowData(data_all) |> as_tibble(),
    {
      .
    },
    by ="ID"
  ) 

fit48 <- lmFit((assays(data_48h)$loess), design = design48h)
fit48 <- eBayes(fit = fit48)
Results$randpcs_int48 <- topTable(fit48, p.value = 1, number = Inf, coef = "rand_codXshock") %>%
  rownames_to_column(., var = "ID") %>%
 # arrange(padj) %>%
  right_join(rowData(data_all) |> as_tibble(),
    {
      .
    },
    by ="ID"
  ) 

fit72 <- lmFit((assays(data_72h)$loess), design = design72h)
fit72 <- eBayes(fit = fit72)
Results$randpcs_int72 <- topTable(fit72, p.value = 1, number = Inf, coef = "rand_codXshock") %>%
  rownames_to_column(., var = "ID") %>%
 # arrange(padj) %>%
  right_join(rowData(data_all) |> as_tibble(),
    {
      .
    },
    by ="ID"
  ) 
#saveRDS(Results, "../NBIS_analyses/Results/DEResults.rds")
```


```{r, include=FALSE, include=TRUE}
Results <- readRDS("../NBIS_analyses/Results/DEResults.rds")
sig1 <- lapply(Results, function(x) {dplyr::filter(x, adj.P.Val<0.05)})
#sig1 <- within(sig1, rm(All_rand, All_pcscore))
#sig_rand <- 
sig1List <- unlist(lapply(Results, function(x) {dplyr::filter(x, adj.P.Val < 0.05) %>% rownames(x)}), use.names = FALSE) %>% unique() ## Extract significant proteins at all time points

rowData(data_all)$variance <- assays(data_all)$loess %>% rowVars(na.rm = TRUE)
data_f <- data_all[rownames(data_all) %in% sig1List, ]
```

#### Association with randomization code at 24 h

No significant associations were found at 24h with temperature randomization.

```{r}
#DT::datatable(round(sig1$rand_24,5))
```
#### Association with randomization code at 48 h

Six proteins were associated with temperature randomization at 48h

```{r}
DT::datatable(round(sig1$rand_48,5))
```
#### Association with randomization code at 72 h

No significant associations with temperature randomization were found at 72h.

```{r}
#DT::datatable(round(sig1$rand_72,5))
```


### Heat Plots for randomization code {.tabset}

We selected all proteins significantly regulated with randomization code. In this analyses 6 Proteins associated with randomization code were found only at 48 timepoint.


#### Heatmap of siginificatly associated protein with temperature randomization

```{r}
### Heatmap for significantly DE proteins
sig_rand <- unlist(lapply(sig1[1:3], rownames, usenames)) %>% as.vector()
#intData <- assays(data_24h)$loess[sig_rand,]
intData <- data_24h[sig_rand,]
#png("../Results/Figures/HeatmapInt.png", units = "cm", height = 30, width = 30,res = 300)
pheatmap(assays(intData)$loess, filename = "./Results/Figures/rand_code_full.pdf",
         display_numbers=FALSE, 
         fontsize=7, 
         cluster_cols=TRUE, 
         cluster_rows = FALSE, 
         scale = "row",
         labels_col =  intData$randomisation_code)
#dev.off()
```

#### Heatmap of foldchanges at 3 timepoints

```{r, fig.cap="Heatmap of foldchanges at 3 timepoints"}
fcc <- as.matrix(bind_cols(Results$rand24[sig_rand,]$logFC, 
                           Results$rand_48[sig_rand,]$logFC, 
                           Results$rand72[sig_rand,]$logFC 
                           )
                 )

dimnames(fcc) <- list(sig_rand, c("t1", "t2", "t3") )
pheatmap(fcc, filename = "./Results/Figures/rand_code_foldchage.pdf",
         cluster_cols = FALSE, 
         cellwidth = 12, 
         cellheight = 20)

```

## 2) Results for cpc score phenotype variable tables 

### 2A) Tables {.tabset}

#### Association with randomization code at 24 h 

In total `r sum(Results$pcscore24$adj.P.Val <0.05)` proteins were significantly regulated at 24h timepoint in association with CPC score.

```{r}
DT::datatable(round(head(Results$pcscore24, 10),5))
```

#### Association with randomization code at 48 h

In total `r sum(Results$pcscore48$adj.P.Val <0.05)` proteins were significantly regulated at 48h timepoint in association with PCS score.

```{r}
DT::datatable(round(head(Results$pcscore48, 10),5))
```

#### Association with randomization code at 72 h

In total `r sum(Results$pcscore72$adj.P.Val <0.05)` proteins were significantly regulated at 72h timepoint in association with PCS score.

```{r}
DT::datatable(round(head(Results$pcscore72, 10),5))
```

### 2B) Figures {.tabset}

#### Heatmap of significantly associated proteins at all three time-points

```{r}
### Heatmap for significantly DE proteins
sig_pcs <- unlist(lapply(sig1[4:6], rownames, usenames)) %>% as.vector() %>% unique()
#intData <- assays(data_24h)$loess[sig_rand,]
intData <- data_24h[sig_pcs,]

#png("../Results/Figures/HeatmapInt.png", units = "cm", height = 30, width = 30,res = 300)
pheatmap(assays(intData)$loess, filename = "./Results/Figures/cpc_heatmap_full.pdf",
         display_numbers=FALSE, 
         fontsize=7, 
         cluster_cols=TRUE, 
         cluster_rows = FALSE,
         scale = "row",
         labels_col =  intData$binarysecout180daycpcscore)
#dev.off()
```

#### Heatmap of Foldchanges of significantly associated proteins at three timepoints 

```{r}
fcc_pcs <- as.matrix(bind_cols(Results$pcscore24[sig_pcs,]$logFC, 
                           Results$pcscore48[sig_pcs,]$logFC, 
                           Results$pcscore72[sig_pcs,]$logFC 
                           )
                 )

dimnames(fcc_pcs) <- list(sig_pcs, c("t1", "t2", "t3") )

pheatmap(fcc_pcs, filename = "./Results/Figures/cpcheatmap_foldchanges.pdf", 
         cluster_cols = FALSE, 
         cellwidth = 12, 
         cellheight = 9)

```

### 2C) PCS * Temperature interaction {.tabset}

> Would it be
possible to switch these colours so that the *poor outcome is in red andn good outcome is in blue*? We have also agreed that the accepted terms for
the CPC-outcome are *good outcome and poor outcome*.
*Instead of 0 and 1 on the x-axis, would it be possible to use 36°C (1) and 33°C (0)?* and *change the label of the x-axis to temperature treatment?*.
>


#### Interaction at 48 hours p.adj = 0.17, P = 0.0004294011
```{r, fig.cap= "Boxplot indicating differential effect of temeprature randomization in different CPC score categories on protein P08294 intensity"}
intData1 <- assays(data_24h)$loess["P08294",]
tDatat <- cbind(as.data.frame(colData(data_24h)),intData1)

# Change the position
p <-ggplot(tDatat, aes(x = as.factor(randomisation_code), y = intData1, fill= as.factor(binarysecout180daycpcscore))) +
  geom_boxplot(position=position_dodge(1)) +
  scale_fill_discrete(name="180 day\nCPC Score",
                         breaks=c("0", "1"),
                         labels=c("Functional", "Impaired"))+
  labs(title="CPC Score*Temperature interaction for\nExtracellular superoxide dismutase(P08294) at 48h", x="Temperature randomization Category" , y= expression("Log"[2]~"protein intensity"))

p <- p + geom_dotplot(binaxis='y', stackdir='center',
                 position=position_dodge(1))

ggsave(p, filename = "./Results/Figures/CPC ScoreXTemperature.pdf")


```

#### Interaction at 3 timepoints

```{r, eval=TRUE, include=TRUE}
data_all <- readRDS("../data_sumExpfilImp.rds")
#data_all <- readRDS("../normalized_filtered_se.rds")
data_all_f <- data_all[, (data_all$shockable=="0" | data_all$shockable==1) & !is.na(data_all$shockable)]
rowData(data_all)$variance <- assays(data_all)$loess %>% rowVars(na.rm = TRUE)
#data_all <- data_all[rowData(data_all)$variance>0.2 | rowData(data_all)$variance<5, ]
#saveRDS(data_all_f,"../data_all_f.rds")
data_24h <- data_all_f[, data_all_f$time=="-24-"]
data_48h <- data_all_f[, data_all_f$time=="-48-"]
data_72h <- data_all_f[, data_all_f$time=="-72-"]
```

```{r, warning=FALSE}
intData2 <- assays(data_all_f)$loess["P08294",]
tDatat2 <- cbind(as.data.frame(colData(data_all_f)),intData2)

p<-ggplot(tDatat2, aes(x = as.factor(randomisation_code), y = intData2, fill= as.factor(binarysecout180daycpcscore))) +
  geom_boxplot(position=position_dodge(1)) +
  scale_fill_discrete(name="180 day\nPCS Score",
                         breaks=c("0", "1"),
                         labels=c("Functional", "Impaired"))+
  labs(title="PCS Score*Temperature interaction for\nExtracellular superoxide dismutase(P08294) at 3 time points", x="Temperature randomization Category" , y= expression("Log"[2]~"protein intensity")) + facet_grid(~time)


p <- p +geom_dotplot(binaxis='y', stackdir='center',
                 position=position_dodge(1))

ggsave(p, filename = "./Results/Figures/CPC ScoreXTemperature_timePoints.pdf")
```


##3) Shockable vs.non Shockable differential abundance analyses



```{r, include=FALSE, eval=FALSE}


#designAll <- model.matrix(~ 1 + data_all$randomisation_code)
design24h <- model.matrix(~ 1 + data_24h$shockable)
design48h <- model.matrix(~ 1 + data_48h$shockable)
design72h <- model.matrix(~ 1 + data_72h$shockable)

#colnames(designAll) <- c("(Intercept)", "rand_cod")
colnames(design24h) <- c("(Intercept)", "shock")
colnames(design48h) <- c("(Intercept)", "shock")
colnames(design72h) <- c("(Intercept)", "shock")


Results_shock <- list()

# fitAll <- lmFit((assays(data_all)$loess), design = designAll)
# fitAll <- eBayes(fit = fitAll)
# Results$All_rand <- topTable(fitAll, p.value = 1, number = Inf, coef = "rand_cod")

fit24<- lmFit((assays(data_24h)$loess), design = design24h)
fit24 <- eBayes(fit = fit24)
Results_shock$shock24 <- topTable(fit24, p.value = 1, number = Inf, coef = "shock")

fit48 <- lmFit((assays(data_48h)$loess), design = design48h)
fit48 <- eBayes(fit = fit48)
Results_shock$shock48 <- topTable(fit48, p.value = 1, number = Inf, coef = "shock")

fit72 <- lmFit((assays(data_72h)$loess), design = design72h)
fit72 <- eBayes(fit = fit72)
Results_shock$shock72 <- topTable(fit72, p.value = 1, number = Inf, coef = "shock")
#saveRDS(Results_shock, "../NBIS_analyses/Results/Results_shock.rds")
```

### Tables{.tabset}

```{r}
Results_shock <- readRDS("../NBIS_analyses/Results/Results_shock.rds")
```

#### 24h

In total `r sum(Results_shock$shock24$adj.P.Val <0.05)` proteins were significantly regulated at 24h timepoint in association with Shock variable.

```{r}
Results_shock <- readRDS("../NBIS_analyses/Results/Results_shock.rds")
DT::datatable(round(head(Results_shock$shock24, 10),5))
```


#### 48h

In total `r sum(Results_shock$shock48$adj.P.Val <0.05)` proteins were significantly regulated at 48h timepoint in association with Shock variable.


```{r}
DT::datatable(round(head(Results_shock$shock48, 10),5))
```

#### 72h

In total `r sum(Results_shock$shock72$adj.P.Val <0.05)` proteins were significantly regulated at 72h timepoint in association with Shock variable.


```{r}
DT::datatable(round(head(Results_shock$shock72, 10),5))
```


### Figures{.tabset}

#### Heatmap of proteins at 24h

```{r, fig.cap="Heat map of Protein instensities in all samples"}
### Heatmap for significantly DE proteins
sigs <- lapply(Results_shock, function(x) {dplyr::filter(x, adj.P.Val<0.05)})
sig_shock <- unlist(lapply(sigs, rownames, usenames)) %>% as.vector() %>% unique()
#intData <- assays(data_24h)$loess[sig_rand,]
intDatas <- data_24h[sig_shock,]
intDatas
#png("../Results/Figures/HeatmapInt.png", units = "cm", height = 30, width = 30,res = 300)
pheatmap(assays(intDatas)$loess, filename = "./Results/Figures/Shockheatmap_full.pdf",
         display_numbers=FALSE, 
         fontsize=7, 
         cluster_cols=TRUE, 
         cluster_rows = TRUE,
         scale = "row",
         labels_col =  intDatas$shockable)

```

#### Heatmap of fold changes at 3 timepoints

```{r, fig.cap="Heat map of Fold Changes at 3 timepoints"}
fcc_shock <- as.matrix(bind_cols(Results_shock$shock24[sig_shock,]$logFC, 
                           Results_shock$shock48[sig_shock,]$logFC, 
                           Results_shock$shock72[sig_shock,]$logFC 
                           )
                 )

dimnames(fcc_shock) <- list(sig_pcs, c("t1", "t2", "t3") )
pheatmap(fcc_shock, filename = "./Results/Figures/Shockheatmap_foldchanges.pdf",
         cluster_cols = FALSE, 
         cellwidth = 12, 
         cellheight = 9)


```

```{r}
sessionInfo()
```

