---
title: "Change_plots"
format: html
editor: visual
---

## Quarto


```{r}
library(SummarizedExperiment)
library(tidyverse)
library(ggplot2)
```



```{r, echo=FALSE, include=FALSE}
data_all <- readRDS("../data_sumExpfilImp.rds")
#data_all <- readRDS("../normalized_filtered_se.rds")
data_all <- data_all[, !is.na(data_all$randomisation_code)]
rowData(data_all)$variance <- assays(data_all)$loess %>% rowVars(na.rm = TRUE)
#data_all <- data_all[rowData(data_all)$variance>0.2 | rowData(data_all)$variance<5, ]
data_24h <- data_all[, data_all$time=="-24-"]
data_48h <- data_all[, data_all$time=="-48-"]
data_72h <- data_all[, data_all$time=="-72-"]
#data <- list(loess_data= assays(data_all)$loess %>% as.data.frame() %>% rownames_to_column(., var = "Protein.Group") , sample_data=colData(data_all) %>% as.data.frame() %>% rownames_to_column(., var = "Sample_ID"), proetin_anno=as.data.frame(rowData(data_all)))
#writexl::write_xlsx(x = data , "./Results/Data.xlsx")
```


### 2C) PCS * Temperature interaction {.tabset}

> Would it be
possible to switch these colours so that the *poor outcome is in red andn good outcome is in blue*? We have also agreed that the accepted terms for
the CPC-outcome are *good outcome and poor outcome*.
*Instead of 0 and 1 on the x-axis, would it be possible to use 36°C (1) and 33°C (0)?* and *change the label of the x-axis to temperature treatment?*.
>


#### Interaction at 48 hours p.adj = 0.17, P = 0.0004294011
```{r, fig.cap= "Boxplot indicating differential effect of temeprature randomization in different CPC score categories on protein P08294 intensity"}
intData1 <- assays(data_24h)$loess["P08294",]
tDatat <- cbind(as.data.frame(colData(data_24h)),intData1)

# Change the position
p <- ggplot(tDatat, aes(x = as.factor(randomisation_code), y = intData1, fill= as.factor(binarysecout180daycpcscore))) +
  geom_boxplot(position=position_dodge(1)) +
  scale_fill_manual(name="CPC outcome",
                         breaks=c("0", "1"),
                         labels=c("Good outcome", "Poor outcome" ),
                    values=c("darkblue", "darkred"))  +
  scale_x_discrete(breaks=c("0","1"),
        labels=c("33°C", "36°C"))+ 
  labs(title="CPC outcome by Temperature interaction plot for\nExtracellular superoxide dismutase(P08294) at 48h", x="Temperature treatment" , y= expression("Log"[2]~"protein intensity")) 
p <- p + geom_dotplot(binaxis='y', stackdir='center',
                 position=position_dodge(1))

p
ggsave(p, filename = "./Results/Figures/CPC_ScoreXTemperature_modified.png")


```

```{r, warning=FALSE}
intData2 <- assays(data_all)$loess["P08294",]
tDatat2 <- cbind(as.data.frame(colData(data_all)),intData2)


p2 <- ggplot(tDatat2, aes(x = as.factor(randomisation_code), y = intData2, fill= as.factor(binarysecout180daycpcscore))) +
  geom_boxplot(position=position_dodge(1)) +
  scale_fill_manual(name="CPC outcome",
                         breaks=c("0", "1"),
                         labels=c("Good outcome", "Poor outcome" ),
                    values=c("darkblue", "darkred"))  +
  scale_x_discrete(breaks=c("0","1"),
        labels=c("33°C", "36°C"))+ 
  labs(title="CPC outcome by Temperature interaction plot for\nExtracellular superoxide dismutase(P08294) at 3 time points", x="Temperature treatment" , y= expression("Log"[2]~"protein intensity")) + facet_grid(~time)


p2 <- p2 +geom_dotplot(binaxis='y', stackdir='center',
                 position=position_dodge(1))

p2
ggsave(p2, filename = "./Results/Figures/CPC ScoreXTemperature_timePoints_modified.png", width = 9)
```