---
title: "Materials and Methods"
author: "Ashfaq Ali"
date: '2022-04-07'
output: html_document
bibliography: packages.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## QC and statistical analyses

> Data QC: Sample ID *SEHBG0019* missing in the clinical data
 - Samples IDs *SEMAL* in proteomics while *SEMMO* in the clinical data
 
### Data filtering 

Data was analysed for patterns in missingness using `mice`package in R to identify any samples of proteins. 
*Missing values:* Proteins missing in more than 30% of the samples were filtered out.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(SummarizedExperiment)
library("tidyverse")
sumExpObj <- readRDS( "../rawaDataSe.rds")
data_all <- readRDS("../data_sumExpfilImp.rds")
data_all <- data_all[, !is.na(data_all$randomisation_code)]
rowData(data_all)$variance <- assays(data_all)$loess %>% rowVars(na.rm = TRUE)
#data_all <- data_all[rowData(data_all)$variance>0.2 | rowData(data_all)$variance<5, ]
data_24h <- data_all[, data_all$time=="-24-"]
data_48h <- data_all[, data_all$time=="-48-"]
data_72h <- data_all[, data_all$time=="-72-"]
```

Initial data had `r dim(sumExpObj)[1]` proteins and `r dim(sumExpObj)[2]` samples which was filtered to `r `dim(data_all)[1]` proteins and `r dim(data_all)[2]` samples.
Out of 210, 7 samples had NA values for randomization code variable and were exluded from analyses.

### Normalization 
Data was normalised using `limma::normalizeCyclicLoess()` function provided by Limma package in R. 

### Exploratory PCA analyses

Principle component analyses was used to perform exploratory analyses of the proteomics data. Explained variance by each principle component was calculated using the output from prcomp´ function in R. Correlation between principle components and phenotype (clinical) variables was calculated using ´lm´ function in R and "adj.r.squared" values were extracted from the summary statistics produced by the ´lm´ function.

## Statistical analyses

The likely differentially abundant proteins were extracted by using `topTable()`function. We used the "BH" (aka "fdr") and "BY" methods of Benjamini, Hochberg, and Yekutieli control the false discovery rate, the expected proportion of false discoveries among the rejected hypotheses. 

## Variables and models

In general, the data from three time points, 24h with `r dim(data_24h)[2]`, 48h `r dim(data_48h)[2]` and 72 `r dim(data_72h)[2]`hours were analyzed separately for uni variate statistics. And the interaction term was included in the model formula wherever applicable.

### Randomization code (0 0r 1):

Positive change indicates upregulation in 1 vs 0. Negative fold change indicates downregulation in 1.

```{r, warning=FALSE, echo=FALSE}
knitr::kable(rbind(t1 = table(data_24h$randomisation_code), t2= table(data_48h$randomisation_code), t3=table(data_72h$randomisation_code)))
```
### CPC Score (0 or 1): 

Positive change indicates upregulation in 1 vs 0. Negative fold change indicates downregulation in 1. 

```{r, warning=FALSE, echo=FALSE}
knitr::kable(rbind(t1 = table(data_24h$binarysecout180daycpcscore), t2= table(data_48h$binarysecout180daycpcscore), t3=table(data_72h$binarysecout180daycpcscore)))
```

### Shockable vs. non shockable heart:

Positive change indicates upregulation in 1 vs 0. Negative fold change indicates downregulation in 1.
Samples with NA values were filtered out.

```{r, echo=FALSE, warning=FALSE}
knitr::kable(rbind(t1 = table(data_24h$shockable), t2= table(data_48h$shockable), t3=table(data_72h$shockable)))
```

## Dead vs. Alive at 72h:

Positive change indicates upregulation in 1 vs 0. Negative fold change indicates downregulation in 1.

In total there were `r table(data_all$dead72)[2]` dead events at 72 hours. 

## Dead vs. Alive at followup:

Positive change indicates upregulation in 1 vs 0. Negative fold change indicates downregulation in 1. 
in total `r table(data_all$deadatfollowup)[2]` patients had died at followup.



## Interactions

Model formulae for interactions tested was as following. The interpretation of these would based on the boxplots as layed out in the figures.

~ 1 + randomisation_code*binarysecout180daycpcscore

~ 1 + randomisation_code*shockable

## Heatmaps

The significantly differentially abundant protein (adj p < 0.05) at any of the timepoint for respective predictor variable was extracted. Plots of respective fold changes were plotted in a heatmap. Heatmap Plots of individual protein intensities were also provided. 


## References

```{r, echo=FALSE}
packages = c("base","mice", "limma")
knitr::write_bib(packages, "packages.bib")
knitr::write_bib(packages)
```

